\documentclass[sigconf,balance=false]{acmart}
\usepackage{popets}
\usepackage{amsmath}

\setcopyright{popets}
\copyrightyear{YYYY}
\acmYear{YYYY}
\acmVolume{YYYY}
\acmNumber{X}
\acmDOI{XXXXXXX.XXXXXXX}
\acmISBN{}
\acmConference{Proceedings on Privacy Enhancing Technologies}
\settopmatter{printacmref=false,printccs=false,printfolios=true}





 

  
\begin{document}
\author{Alexander Gamero-Garrido}
\affiliation{%
 \institution{University of California, Davis}
 \streetaddress{}
 \city{}
 \state{}
 \country{}
 }

\author{Kicho Yu}
\affiliation{%
 \institution{University of Southern California}
 \streetaddress{}
 \city{}
 \state{}
 \country{}
 }

\author{Sumukh Vasisht Shankar}
\affiliation{%
 \institution{Yale University}
 \streetaddress{}
 \city{}
 \state{}
 \country{}
 }

\author{Sachin Kumar Singh}
\affiliation{%
 \institution{University of Utah}
 \streetaddress{}
 \city{}
 \state{}
 \country{}
 }

\author{Sindhya Balasubramanian}
\affiliation{%
 \institution{Northeastern University}
 \streetaddress{}
 \city{}
 \state{}
 \country{}
 }

\author{Alexander Wilcox}
\affiliation{%
 \institution{Northeastern University}
 \streetaddress{}
 \city{}
 \state{}
 \country{}
 }

\author{David Choffnes}
\affiliation{%
 \institution{Northeastern University}
 \streetaddress{}
 \city{}
 \state{}
 \country{}
 }








  \title{Empirically Measuring Data Localization in the EU}



\begin{abstract}EU data localization regulations limit data transfers to non-EU countries with the GDPR. However, BGP, DNS and other Internet protocols were not designed to enforce jurisdictional constraints, so implementing data localization is challenging. 
Despite initial research on the topic, 
little is known about if or how companies currently operate their server infrastructure to comply with the regulations.
We close this knowledge gap by empirically measuring the extent 
to which servers and routers that process EU requests are located 
outside of the EU (and a handful of ``adequate'' non-EU countries). 
The key challenge is that both browser measurements (to infer relevant endpoints) 
and data-plane measurements (to infer relevant IP addresses) are needed, 
but no large-scale public infrastructure allows both. We build a novel methodology 
that combines BrightData (browser) and RIPE Atlas (data-plane) probes,
 with joint measurements from over 1,000 networks in 20 EU countries.
We find that, on average, 2.2\% of servers %
serving users in each EU country are located in non-adequate destination countries (1.4\% of known trackers). 
Our findings
suggest that data localization policies are largely being followed
by content providers, though there are %
exceptions. %
\end{abstract}

  \keywords{Data Localization, GDPR, Internet Measurement}


 

\maketitle
\input{intro.tex}
\input{method-domains.tex}
\input{method-geolocation.tex}

\input{results-overview.tex}
\input{related.tex}

{\sloppy
\scriptsize{
\bibliographystyle{ACM-Reference-Format}
\interlinepenalty=10000
\bibliography{references}}}

\input{appendix.tex}

\end{document}
